Prospects of detecting gamma-ray emission from galaxy clusters: cosmic rays and dark matter annihilations

### 2011-05-07 paper.tex
abstract commited

\begin{abstract}
... to be written...  if $\mlim \approx 10^{-6}\,\msun$ then the boost
  from SFE is smaller than about a factor 5 in the MW and 2 in Fornax, or
  alternatively if SFE correct then $\mlim > 10^2\,\msun$.
\end{abstract}

>>>>

\begin{abstract}
  We study the possibility for detecting gamma-ray emission in galaxy clusters. We consider 1)
  cosmic ray (CR) induced pion decay which is thought to dominate the astrophysical signal from
  clusters, 2) different representative benchmark models of supersymmetric DM, and 3) leptophilic
  models of dark matter (DM) annihilation that recently have been proposed to explain the positron
  excess seen by PAMELA and include a Sommerfeld enhancement of the particle physics cross section
  over its standard value. To model DM annihilation, we consider continuum emission as a result of
  the hadronization the annihilating neutralinos, internal bremsstrahlung, and inverse Compton
  emission from the cosmic microwave background as well as a realistic spatial and spectral
  distribution of dust and stellar light.  After the galactic center which hosts many astrophysical
  sources, galaxy clusters are the best known target class to detect DM annihilation. We predict
  Virgo and Fornax to be the brightest DM sources and find a particularly low CR induced background
  for Fornax. Because for a given minimum substructure mass as given by the DM free-streaming scale,
  galaxy clusters are the largest collapsed objects which maximizes the boost of the smoothly
  distributed annihilation flux by a factor of more than 1000. Since the annihilation flux of
  substructures is mostly contributed by the regions around the virial radius, the resulting surface
  brightness profiles are almost flat. This makes it very challenging to detect this flux with
  imaging air Cherenkov telescopes since their sensitivity drops linearly with radius and they
  typically encounter have $5-10$ linear resolution elements across a cluster. Assuming cold dark
  matter with a substructure mass distribution down to an Earth mass, we rule out the leptophilic
  models in present form in 14 clusters, and limit the boost from Sommerfeld enhancement in Fornax
  to be less than 2 which corresponds to a value in the Milky Way of less than 5. Alternatively, if
  the Sommerfeld enhancement factor is realized in Nature, this would imply limiting substructure
  mass in Fornax of $\mlim > 10^2\,\msun$ corresponding to a Milky Way value of $\mlim > 5\times
  10^3\,\msun$ which would be evidence for an unknown form of warm dark matter. We find that Fermi
  three-year data will be able to constrain the smallest mass of substructures using the
  supersymmetric benchmark models or alternatively their cross sections for a given substructure
  boost. Although the Perseus and Ophiuchus clusters are expected to have the largest CR induced
  flux according to our semi-analytic model that is derived from cosmological hydrodynamical cluster
  simulations, their central active galactic nucleus as well as the Galactic foreground makes it
  hard to detect this flux in the Fermi band. The next to best targets to detect the CR flux are the
  Virgo, Coma, and Norma clusters -- the Fermi upper limits are closing in on our predictions and
  will soon start to constraining the underlying CR physics such as shock acceleration efficiencies
  or CR transport properties.
\end{abstract}

\del{
--) the flux from annihilating DM is boosted by about a factor 1000 in
  clusters
--) flat surface brightness profiles with substructures
--) Fornax is a great target for indirect DM studies because the relative
  low flux from CRs and high flux from DM
--) difficult to detect DM without a Sommerfeld boost with Cherenkov
  telescopes since the boost from substructures is extended while the
  sensitivity of Cherenkov telescopes drops linearly with radius
  outside the point spread function (CHECK)
--) the IC upscattered photons from stars and dust are dominated in the
  relevant energy regimes by the continuum emission for our DM
  benchmark models
--) we provide a rather simple model to calculate the spatial and spectral
  distribution of dust and stars in cluster (might be used for
  galaxies? with different spatial profile)
--) we rule out the leptophilic models in present form in 14 clusters,
  and limit the boost from Sommerfeld enhancement in the MW to less
  than 10
--) the tightest limits on DM and CRs from Fermi are in the 1-10 GeV
  regime, where Fornax, M49, and NGC4636 are the most constraining
  "clusters" for the DM and 
... to be written...  if $\mlim \approx 10^{-6}\,\msun$ then the boost
  from SFE is smaller than about a factor 5 in the MW and 2 in Fornax, or
  alternatively if SFE correct then  $\mlim > 10^2\,\msun$.
}
